### 2015-07-09 narrative/placement.tex

Our proposed research will address data management challenges across the entire 
data life cycle in an application driven and system aware manner. 
%given our thesis that more knowledge, and more specific knowledge about the
%data, can allow the management of data to be optimized with reduced cost.
Some of the high level research questions that we will explore are as follows: 
(1) How can we initially place data so that it can be discovered and consumed efficiently?,
%What methods can be utilized in the initial placement of data that do not impose expensive
%costs on discovery, consumption and migration? 
(2) How can the placement and migration of data across a multi-tiered storage hierarchy be 
optimized at runtime, both from the application and system perspective?, {\color{red} where is 3?} 
%(3) How can we migrate data throughout the layers of storage
%without making the discovery process unbounded? 
(4) How can knowledge about the application used to better prepare the data for consumption?, and 
(5) When and how do we make the decision to purge data?  %using this additional knowledge?

@@

\subsubsection{Data Placement and Movement}
\label{sec:init-plac-data}

\paragraph{Background:}
When an application outputs or accesses data the storage and middleware layers needs to
decide \textit{what} data is placed \textit{where} in the multi-level storage system. 
This placement decision can have a significant impact on data management 
throughout the life time of the data. For example, our past work on data 
staging~\cite{tongipdps15,qiansc15,docan2012dataspaces,abbasi2010datastager} 
on HPC systems with multi-level memory structures has shown that different 
output techniques targeting different layers of the storage hierarchy can have a 
significant impact on the overhead observed by the application for I/O operations.

@@

A key requirement is application-driven runtime mechanisms for dynamically managing 
data placement across the layers of the distributed storage hierarchy throughout the 
data lifecycle, coordinating data movement and data sharing between the components 
of the application workflow, with the overarching goal of maximizing the relative utility 
to the application as well as the system while reducing access costs. 
%
As noted before, the complexities of heterogeneous multi-level storage structures requires 
adaptive placement policies are required to be implemented to optimally utilize storage 
resources vertically (across deep memory hierarchies) and horizontally (across nodes 
within a memory level) while accommodating dynamic application requirements and transient 
system states. 
%
Our approach in the proposal is to increase knowledge about the data and its use within the 
application, and leverage this knowledge to drive data placement and overall management.

@@

\paragraph{ Runtime tracking and estimation:}
Autonomic data placement and/or movement will leverage the information gathered 
at runtime together with predictive analysis requirements. This information may include 
runtime spatial-temporal data access patterns, and/or physical network topology
information, which will be tracked at runtime, and then be used to estimate and predict 
the data access behaviors of application workflow at runtime. 
A key research aspect here will be to anticipate accesses, for example using learning techniques or using
application knowledge, and use this information to prefetch data objects.
For example, we will identify and catalog key phases and access patterns
in the target workflows and then use it to characterize patterns at runtime
and trigger appropriate placement and/or movement actions. Our initial 
work explore these ideas is presented in~\cite{qiansc15}.

@@

\subsubsection{Migration}
\label{sec:migration}

\paragraph{Background:} 
A complementary aspects of the data lifecycle management is the management and 
%Once the system has completed the initial data placement with minimal
%overhead to the application, it is up to the storage system to manage the
migration of data across storage servers and across the hierarchy. 
An important research aspect here is how a liberated Hierarchical Storage
Manager (HSM) designed for on-line use, such as Sirocco, might be employed in
providing support for the migration of data. Sirocco autonomously groups the
distributed set of media and service nodes into like groups and manages the
movement of data between these groups. The grouping is managed using
attributes such as latency to first byte, bandwidth to and from the media, and
some function representing the resilience capabilities of the attached media.
In addition to these static attributes, when grouping servers into
pseudo-tiers, Sirocco also may employ more dynamic attributes such as CPU load
and the rate-of-change of the media use.

Grouping into pseudo-tiers is not a formal, nor rigid, concept within
Sirocco, at present. In classic HSM the tiering is often a formally
expressed concept, embodied both in the architecture and the implementation.
In Sirocco, though, it is not directly expressed. Instead, a client or
server that attempts to make a decision relative to migration and staging
uses these attributes to constrain the choice of candidates. For instance, a
server acting as an off-node writeback cache for a compute node might only
have access to volatile dynamic RAM for use as storage media and a client
might deposit data that is marked for persistent media as a resilience
constraint. At some point, then, such a server must copy or move the data to
another Sirocco server with media matching, or exceeding, the resilience
constraint. At that point in time, all servers managing appropriate
persistent media are candidates but second-order attributes, such as latency
and bandwidth of a candidate's managed media would tend to dominate the
ranking. Thus, the concept of a ``tier'' becomes an artifact of decision
rather than a formal concept.

@@

\paragraph{Migration Policies:}
The policies that govern the migration of data across groups and across
tiers have similar challenges as data placement. The migration
policy must encapsulate, at a high level, the constraints defined by the
application, but must also manage resources on each server for capacity and
load reasons. We will explore policy formulations that allow \Sir to move
data within a tier to continue to meet service objectives. Another component
of migration is the temporal function that determines the utility of data
through its life time on storage. If they utility decays over time, for
example, the initial constraints of data (such as resilience requirements)
will be modified and allow more leeway in how the system handles migration.
However, as data is widely distributed the scalability of this approach must
be investigated. 
We will explore how the utility function can be used to modify migration
parameters in a scalable manner, without requiring a centralized authority
to make decisions.

\paragraph{Bounded estimates for lookup:}
One challenge for storage systems with continuous decentralized migration is
the timeliness of lookup when applications issue read calls. This is due to
how no central directory exists for where data is at anytime. There are
solutions to this problem that utilize lookup tables which are updated
whenever a server initiates a migration process. However, these solutions
are not scalable. In order to have predictable performance the lookup
process must complete in a bounded time. Another solution here is to use a
distributed hashing function to determine exactly where data is placed,
reducing lookup to an $O(1)$ operation. We explore the discovery of data in
\S~\ref{sec:naming-discovery}. To support this discovery process, we will
investigate how the migration of data can be performed in a way that enables
bounded estimates about time to first byte for a potential read operation.

@@

\subsubsection*{Making data available on multiple tiers}
\label{sec:manag-data-mult}

\paragraph{Background:} Providing data to users for knowledge discovery is
one of the primary purposes of our storage system design. There is a
spectrum of approaches on how
this can be accomplished: On one end the storage system can aim to maintain
full resolution data in the highest tiers of the storage hierarchy. This
approach is, obviously, limited due to the disparity between size of data
and available space in the higher tiers. Another approach is to utilize the
techniques described in \S\ref{sec:data-refactor} to create auxiliary views of
the data that, with user defined error, represent the data, and place these
smaller data sets on the higher tiers for faster access.

@@

\paragraph{Locality Properties:} The decision on when to thin and when to
enrich data on a particular level depends on \emph{access patterns}.
Approaches that leverage access patterns in applications at leadership facilities
 have shown to greatly improve performance and reduce the amount
of overhead required for data management~\cite{he:hpdc13}. The
challenge is for applications to communicate these access patterns
to the I/O stack layers that manages tiers outside of the address
space. In the proposed work we will develop an abstraction for
\emph{locality} that allow applications or runtime profilers to
describe locality properties of access patterns, and implement
services that leverage locality properties to dynamically determine
the value of thinning or enriching data at a particular tier.
Consider the example of a road navigation system: the locality
properties of each navigation client are based on physical constraints
of speed, direction, likely resolution of map, and the fact that
clients almost never go off-road.

\subsubsection{Purging}
\label{sec:purging-data}
An important phase of the data lifecycle is the purging of data from the system, 
which presents significant challenges, both from the systems and policy perspectives. 
Deciding what data to purge and when to purge it can be non-trivial, and it is 
essential that this is done with the application in the loop. One approach we will 
explore in this efforts will be extending the notion of system and application utility. 
The former can be used to determine when data object need to purged while the 
latter can help determine the relative importance of data objects to the 
application and identify which data objects can be purged. Developing policies 
for purging is a more challenging problem and will required engaging with the 
user community. However, a key research problem that will be addressed in
 this effort is developing the abstractions that can be used to consistently and 
 unambiguously express these policies as well as the mechanisms that can be 
 used to correctly enforce these policies.

>>>>

Our research will address data management challenges across the entire 
data life cycle in an application driven and system aware manner. 
%given our thesis that more knowledge, and more specific knowledge about the
%data, can allow the management of data to be optimized with reduced cost.
Some of the high level research questions that we will explore are as follows: 
(1) How can we initially place data so that it can be discovered and consumed efficiently?,
%What methods can be utilized in the initial placement of data that do not impose expensive
%costs on discovery, consumption and migration? 
(2) How can the placement and migration of data across a multi-tiered storage hierarchy be 
optimized at runtime, both from the application and system perspective?,
%(3) How can we migrate data throughout the layers of storage
%without making the discovery process unbounded? 
(3) How can knowledge about the application used to better prepare the data for consumption?, and 
(4) When and how do we make the decision to purge data?  %using this additional knowledge?

@@

\subsubsection{Data Placement and Movement}
\label{sec:init-plac-data}
When an application outputs or accesses data the storage and middleware layers needs to
decide \textit{what} data is placed \textit{where} in the multi-level storage system. 
This placement decision can have a significant impact on data management 
throughout the life time of the data. For example, our past work on data 
staging~\cite{tongipdps15,qiansc15,docan2012dataspaces,abbasi2010datastager} 
on HPC systems with multi-level memory structures has shown that different 
output techniques targeting different layers of the storage hierarchy can have a 
significant impact on the overhead observed by the application for I/O operations.

@@

A key requirement is application-driven runtime mechanisms for dynamically managing 
data placement across the layers of the distributed storage hierarchy throughout the 
data lifecycle, coordinating data movement and data sharing between the components 
of the application workflow, with the overarching goal of maximizing the relative utility 
to the application as well as the system while reducing access costs. 
%
As noted before, the complexities of heterogeneous multi-level storage structures requires 
adaptive placement policies are required to be implemented to optimally utilize storage 
resources vertically (across deep memory hierarchies) and horizontally (across nodes 
within a memory level) while accommodating dynamic application requirements and transient 
system states. 
%
Our approach is to increase knowledge about the data and its use within the 
application, and leverage this knowledge to drive data placement and overall management.

@@

\paragraph{ Runtime tracking and estimation:}
Autonomic data placement and/or movement will leverage the information gathered 
at runtime together with predictive analysis requirements. This information may include 
runtime spatial-temporal data access patterns, and/or physical network topology
information, which will be tracked at runtime, and then be used to estimate and predict 
the data access behaviors of application workflow at runtime. 
A key research aspect here will be to anticipate accesses, for example using learning techniques or using
application knowledge, and use this information to prefetch data objects.
For example, we will identify and catalog key phases and access patterns
in the target workflows and then use it to characterize patterns at runtime
and trigger appropriate placement and/or movement actions. Our initial 
work explore these ideas is presented in~\cite{qiansc15,choi2012mining}.

@@

\subsubsection{Migration}
\label{sec:migration}
A complementary aspects of the data lifecycle management is the management and 
%Once the system has completed the initial data placement with minimal
%overhead to the application, it is up to the storage system to manage the
migration of data across storage servers and across the hierarchy. 
An important research aspect here is how a liberated Hierarchical Storage
Manager (HSM) designed for on-line use, such as Sirocco, might be employed in
providing support for the migration of data. Sirocco autonomously groups the
distributed set of media and service nodes into like groups and manages the
movement of data between these groups. The grouping is managed using
attributes such as latency to first byte, bandwidth to and from the media, and
some function representing the resilience capabilities of the attached media.
In addition to these static attributes, when grouping servers into
pseudo-tiers, Sirocco also may employ more dynamic attributes such as CPU load
and the rate-of-change of the media use.

Grouping into pseudo-tiers is not a formal, nor rigid, concept within
Sirocco, at present. In classic HSM the tiering is often a formally
expressed concept, embodied both in the architecture and the implementation.
In Sirocco, though, it is not directly expressed; instead, a client or
server that attempts to make a decision relative to migration and staging
uses these attributes to constrain the choice of candidates. For instance, a
server acting as an off-node writeback cache for a compute node might only
have access to volatile dynamic RAM for use as storage media and a client
might deposit data that is marked for persistent media as a resilience
constraint. At some point, then, such a server must copy or move the data to
another Sirocco server with media matching, or exceeding, the resilience
constraint. At that point in time, all servers managing appropriate
persistent media are candidates but second-order attributes, such as latency
and bandwidth of a candidate's managed media would tend to dominate the
ranking. Thus, the concept of a ``tier'' becomes an artifact of decision
rather than a formal concept.

@@

\paragraph{Migration Policies:}
The policies that govern the migration of data across groups and across
tiers have similar challenges as data placement. The migration
policy must encapsulate, at a high level, the constraints defined by the
application, but must also manage resources on each server for capacity and
load reasons. We will explore policy formulations that allow \Sir~to move
data within a tier and continue to meet service objectives. Another component
of migration is the temporal function that determines the utility of data
through its life time on storage. If the utility decays over time, for
example, the initial constraints of data (such as resilience requirements)
will be modified and allow more leeway in how the system will handle migration.
However, as data is widely distributed {\it the scalability of this approach must
be investigated}. 
We will explore how the utility function can be used to modify migration
parameters in a scalable manner, without requiring a centralized authority
to make decisions.

\paragraph{Bounded estimates for lookup:}
One challenge for storage systems with continuous decentralized migration is
the timeliness of lookup when applications issue read calls. This is due to
how no central directory exists for where data is at anytime. There are
solutions to this problem that utilize lookup tables which are updated
whenever a server initiates a migration process. However, these solutions
are not scalable. In order to have predictable performance the lookup
process must complete in a bounded time. Our potential solution here is to use a
distributed hashing function to determine exactly where data is placed,
reducing lookup to an $O(1)$ operation. We explore the discovery of data in
\S~\ref{sec:naming-discovery}. To support this discovery process, we will
investigate how the migration of data can be performed in a way that enables
bounded estimates  for a potential read operation.

@@

\subsubsection*{Making data available on multiple tiers}
\label{sec:manag-data-mult}
 Providing data to users for knowledge discovery is
one of the primary purposes of our storage system design. There is a
spectrum of approaches on how
this can be accomplished: On one end the storage system can aim to maintain
fully accurate data in the highest tiers of the storage hierarchy. This
approach is, obviously, limited due to the disparity between size of data
and available space in the higher tiers. Another approach is to utilize the
techniques described in \S\ref{sec:data-refactor} to create auxiliary views of
the data that, with user defined error, represent the data, and place these
high priority  data sets on the higher tiers for faster access.

@@

\paragraph{Locality Properties:} The decision on when to thin and when to
enrich data on a particular level depends on \emph{access patterns}.
Approaches that leverage access patterns in applications at leadership facilities
 have shown to greatly improve performance and reduce the amount
of overhead required for data management~\cite{he:hpdc13}. The
challenge is for applications to communicate these access patterns
to the I/O stack layers that manages tiers outside of the address
space. In the proposed work we will develop an abstraction for
\emph{locality} that allow applications or runtime profilers to
describe locality properties of access patterns, and implement
services that leverage locality properties to dynamically determine
the value of thinning or enriching data at a particular tier.
%Consider the example of a road navigation system: the locality
%properties of each navigation client are based on physical constraints
%of speed, direction, likely resolution of map, and the fact that
%clients almost never go off-road.

\subsubsection*{Purging}
\label{sec:purging-data}
An important phase of the data lifecycle is the purging of data from the system, 
which presents significant challenges, both from the systems and policy perspectives. 
Deciding what data to purge and when to purge it can be non-trivial, and it is 
essential that this is done with the application in the loop. One approach we will 
explore in this efforts will be extending the notion of system and application utility. 
The former can be used to determine when data object need to purged while the 
latter can help determine the relative importance of data objects to the 
application and identify which data objects can be purged. Developing policies 
for purging is a more challenging problem and will require engaging  the 
user community. However, a key research problem that will be addressed in
 this effort is developing the abstractions that can be used to consistently and 
 unambiguously express these policies as well as the mechanisms that can be 
 used to correctly enforce these policies.
